---
title: "Spotify and Billboard Data"
author: "Lucero Mejia, Jason Yan, Shea Frantz, Schawnery Lin"
date: "February 22, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("knitr")
source("analysis.R")
year <- 1997
song_name <- "I Believe I Can Fly"
artist_name <- "R. Kelly"
```

## Section 2: Billboard Data

This section focuses on data collected by Billboard, a collection of the most popular U.S songs from 1950 to 2015, as well as their corresponding data (e.g. track name, popularity and duration) from the Spotify data base. 

### First data set: List of Billboard songs

This specific data set was collected by Juhee Lee, Yinan Wen, and Kevin Schaich, and can be found [here](https://github.com/kevinschaich/billboard). However, the basis of that data set are the actual Billboard rankings for a given year, which can be found [here](https://www.billboard.com/charts).

Sample of Billboards Top 100 data set in `r year`:

``` {r billboard data, echo = FALSE}
billboard <- head(get_year(year))
kable(billboard, col.names = c("Song Title", "Artist", "Position", "number of words", "genre"))
```

The table provides:

- **Song title** - The name of the song.
- **Artist** - The person who produced the song.
- **Position** - The songs position (from 1 to 100) on the Billboards Top 100 list in a given year.
- **Number of words** - The number of words, also known as lyrics, the songs contains. 
- **Genre** - The category of artistic, musical, or literary composition characterized by a particular style the song falls in. At times the song can fall into multiple genres which wil be presented as c().


### Second data set: Spotify data for a particular song

This specific data set was collected by Spotify which can be found [here](https://developer.spotify.com/documentation/web-api/reference/search/search/). We will find the Spotify data for each track in the specific `billboard_data` file.

Here is a sample of the data set for the track "`r song_name`" by `r artist_name`:

```{r spotify data, echo = FALSE}
kable(track_ids(paste(song_name, artist_name)))
```

The table provides:

- **Name** - The name of the song.
- **Duration_ms** - How long (time) the song is presented in milliseconds.
- **Popularity** - Popularity is based on the number of streams. The higher the number of streams is, the more popular that song is.
